---
title: "introduction"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The `pipetime` package lets you measure how long your pipeline (`|>`) operations take. It works with the native R pipe and fits naturally into **tidy workflows**.

You can use `time_pipe()` in two ways:

1.  **Time a single step:** wrap `time_pipe()` around a specific `dplyr` operation. Only that step is measured.

2.  **Time a whole pipeline up to a point:** place `time_pipe()` at the end of a pipeline. It will report the total time for all steps up to that point.

Timing messages can be printed to the console or saved to a file.

``` {r}
# devtools::install_github("CyGei/pipetime")
library(pipetime)
library(dplyr)
```

# Timing a Single Step

Wrap `time_pipe()` around a specific operation:
```{r}
data.frame(x = 1:6) |>
  time_pipe(mutate(y = {Sys.sleep(0.2); rnorm(n(), mean = x)}), "random draw") |>
  mutate(y2 = y^2) |>
  summarise(avg_y = mean(y))
```
Only the `mutate(y = …)` step is timed.

# Timing a Whole Pipeline

Place `time_pipe()` on the data at the end (or after multiple steps):
```{r}
data.frame(x = 1:6) |>
  mutate(y = {Sys.sleep(0.2); rnorm(n(), mean = x)}) |> 
  mutate(z = {Sys.sleep(0.2); rnorm(n(), mean = x^2)}) |> 
  summarise(avg_y = mean(y),
            avg_z = mean(z)) |>
  time_pipe("total pipeline time")
```

Here, the timing includes **all previous operations** in the pipeline, giving the total time.

You can place `time_pipe()` at several points in a pipeline to **cumulatively measure total time up to each point**. Each `time_pipe()` marks a checkpoint and records the elapsed time for all operations since the start of the pipeline:
```{r}
data.frame(x = 1) |> 
  # Step 1: small operation
  mutate(a = { Sys.sleep(1); 1}) |> 
  time_pipe("Step 1") |>   # ~1 sec
  
  # Step 2: multiple operations
  mutate(b = { Sys.sleep(1); 1},
         c = { Sys.sleep(1); 1}) |> 
  time_pipe("Step 2") |>   # ~3 sec
  
  # Step 3: final summarise
  summarise(total = sum(a + b + c),
            pause = {Sys.sleep(1); 1}) |> 
  time_pipe("Total Pipeline")  # ~4 sec
```

# Logging to a File

You can save timing logs to a file using the `log_file` argument:
```{r}
log_file <- tempfile(fileext = ".log")

df <- mtcars |>
  mutate(hp2 = {Sys.sleep(0.1); hp * 2}) |>
  time_pipe("Step 1", log_file = log_file, console = FALSE) |>
  mutate(hp3 = {Sys.sleep(0.1); hp * 3}) |>
  time_pipe("Step 2", log_file = log_file, console = FALSE)

# View log
cat(readLines(log_file), sep = "\n")
```
